---
title: "Introduction to R"
output: ioslides_presentation
author: Laura Melissa Guzman and Juan Zuloaga
---

```{r echo = FALSE, message=FALSE}
library(ggplot2)
library(cowplot)
library(knitr)    
library(png)
```

## Overview

This is a series of R, statistics and conservation workshops.

The goals of these workshops are to:

- Introduce R
- Introduce basic concepts of statistics
- Apply the conceps from R and statistics to conservation problems 


## What is R and RStudio?

R is a open-source programming language commonly used for statistics

RStudio is an integrated development environment (IDE) for R -- A nice place to use R


## Why use R?

- R is free!
- R is the most common software used for statistical analyses
- R has an amazing community


## What can I do with R?

- From simple graphs
- Interactive web apps
- Documents/ notebooks
- Spatial analysis
- Statistical analysis 
- Much more!


## R packages

R allows you to use packages (functions developed by others). By including these packages, you can expand the capabilities of R.

```{r echo = FALSE}
packages_path <- "pics/packages.jpg"
include_graphics(packages_path, dpi  = 100)
```


## You can use R to produce really cool graphs!

```{r fig.height=3, fig.width=4, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
ggplot(data = mtcars, aes(x = disp, y = mpg, colour = factor(gear))) + geom_point() + theme_cowplot() + xlab("Displacement (cu.in.)") + ylab("Miles/(US) gallon") 
```

## Shiny apps

[covid tracker](https://shiny.rstudio.com/gallery/covid19-tracker.html)

```{r echo = FALSE}
shiny_path <- "pics/shiny.png"
include_graphics(shiny_path)
```

## RMarkdown

- [Books!](https://bookdown.org/yihui/bookdown/)
- Presentations! (like this one)
- [Documents](http://timelyportfolio.github.io/rCharts_nyt_home_price/)
- Websites

## Simulations



## Functions

R has many built in functions or you can load many more using packages. 

You can also write your own. They are specially useful if you need to repeat the same code multiple times. 

```{r eval=FALSE}

my_function <- function(a, b){
  a*b
}

```


## Statistics 



## Spatial analysis


## Interaction between R and other softwares

- [Rcpp](http://adv-r.had.co.nz/Rcpp.html)
- [Excel](https://readxl.tidyverse.org)

