---
title: "Introduction to R"
output: ioslides_presentation
author: Laura Melissa Guzman and Juan Zuloaga
---

## Overview

This is a series of R, statistics and conservation workshops.

The goals of these workshops are to:

- Introduce R
- Introduce basic concepts of statistics
- Apply the conceps from R and statistics to conservation problems 


## What is R and RStudio?

R is a open-source programming language commonly used for statistics

RStudio is an integrated development environment (IDE) for R -- A nice place to use R


## Why use R?

- R is free!
- R is the most common software used for statistical analyses
- R has an amazing community

## What can I do with R?





## You can use R to produce really cool graphs!

```{r echo = FALSE, message=FALSE}
library(ggplot2)
library(cowplot)
```

```{r fig.height=3, fig.width=4}
ggplot(data = mtcars, aes(x = disp, y = mpg, colour = factor(gear))) + geom_point() + theme_cowplot() + xlab("Displacement (cu.in.)") + ylab("Miles/(US) gallon") 
```